On Extending a Full-Sharing Design with Batched Scheduling

### 2015-04-15 paper.tex
introduction -> begin

@@

>>>>

Tabling~\cite{Chen-96} is an implementation technique that overcomes
some limitations of traditional Prolog systems in dealing with
redundant sub-computations and recursion. Tabling consists of storing
intermediate answers for subgoals so that they can be reused when a
repeated subgoal appears during the resolution process. Tabling has
become a popular and successful technique thanks to the
ground-breaking work in the XSB Prolog system and in particular in the
SLG-WAM engine~\cite{Sagonas-98}, the most successful engine of
XSB. The success of SLG-WAM led to several alternative implementations
that differ in the execution rule, in the data-structures used to
implement tabling, and in the changes to the underlying Prolog
engine. Implementations of tabling are now widely available in systems
like Yap Prolog, B-Prolog, ALS-Prolog, Mercury, Ciao Prolog and more
recently Picat.

Multithreading in Prolog is the ability to concurrently perform
computations, in which each computation runs independently but shares
the program clauses. When multithreading is combined with tabling, we
have the best of both worlds, since we can exploit the combination of
higher procedural control with higher declarative semantics. To the
best of our knowledge, XSB~\cite{Marques-08} and Yap~\cite{Areias-12a}
were the only Prolog systems that were able to combine both
multithreading with tabling. Yap's threads have their own execution
stacks and only share the code area where predicates, records, flags
and other global non-backtrackable data are stored. For tabled
evaluation, a thread views its tables as private but, at the engine
level, Yap has three designs: (i) The {\bf No-Sharing (NS)} design,
where each thread allocates fully private tables for each new tabled
subgoal called during its computation. (ii) {\bf Subgoal-Sharing
  (SS)}, where threads share part of the table space and the (iii)
{\bf Full-Sharing (FS)}, where threads share the complete table
space.

In this paper we discuss a novel approach for solving the problem of
supporting multithreaded batched scheduling in the FS design and we
present a performance analysis comparison between local scheduling
with batched scheduling. Experimental results show that despite the
extra data structures required to extend the FS design with batched
scheduling, the execution time of the FS design with batched
scheduling is still quite competitive when comparing against local
scheduling.

The remainder of the paper is organized as follows. First, we briefly
introduce some background and related work. Then, we describe our
approach for combining batched scheduling with the FS design. Next, we
show the most important implementation details and finally, we discuss
experimental results and we end by outlining some conclusions.

@@

This section introduces some background needed for the following
sections.
